### 2023-01-23 Paper/Paper.Rmd
Initial commit

---
title: "The Impact of Loadshedding on SWIX sectors using DCC-GARCH models"
documentclass: "elsarticle"
Thesis_FP: FALSE
output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5
    fig_height: 3.5
  html_document:
    df_print: paged

Author1: 'Sven Wellmann'
Ref1: "Stellenbosch University, Stellenbosch, South Africa"
Email1: '20850980\@sun.ac.za'

BottomRFooter: "\\footnotesize Page \\thepage"
addtoprule: TRUE
addfootrule: TRUE

margin: 2.3
bottom: 2
top: 2.5
HardSet_layout: TRUE

linenumbers: FALSE

bibliography: Tex/ref.bib
csl: Tex/harvard-stellenbosch-university.csl

RemovePreprintSubmittedTo: TRUE
Journal: "Journal of Finance"

toc: FALSE # Table of Contents

numbersections: TRUE

fontsize: 11pt
linestretch: 1.2
link-citations: TRUE

AddTitle: TRUE

abstract: |
  Abstract
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

pacman::p_load(modelsummary, gt, knitr, kableExtra, tidyverse, lubridate, tinytex, rmarkdown, crypto2, quantmod, MTS, rmgarch, rugarch, tbl2xts, fmxdat, ggthemes, summarytools, vars, tseries, pander, cowplot, qpcR)

list.files('code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))

list.files('../code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
```

<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

# Data  {-}

```{r Figure1,  warning =  FALSE, fig.align = 'center', fig.cap = "Caption Here \\label{Figure1}", fig.ext = 'png', fig.height = 3, fig.width = 6}

```

```{r Figure2, warning =  FALSE, fig.align = 'center', fig.cap = "Caption Here \\label{Figure2}", fig.height = 3, fig.width = 6, dev = 'png'}

```

# Splitting a page

#  Methodology \label{Meth}

# Results

\hfill

# Conclusion

<!-- Make title of bibliography here: -->
<!-- \newpage -->

\newpage

# References {.unnumbered}

::: {#refs}
:::

# Appendix {.unnumbered}

## Appendix A {.unnumbered}

Some appendix information here

## Appendix B {.unnumbered}